---
title: "MA 641 Final Project"
author: "Hrushikesh Thanikonda"
date: "2025-12-12"
output: pdf_document
---
# Appendix 
## Time Series Analysis of Black Rock Daily - Stock Prices (Non- Seasonal)
```{r}
# Importing Blackrock data set 
blk <- read.csv("C:\\Users\\hrush\\OneDrive\\Desktop\\Time Series Analysis Project\\BlackRockData2025.csv")
# calculating returns
n <- length(blk$BLK.Adjusted)
blkReturns <-( blk$BLK.Adjusted[2:n]-blk$BLK.Adjusted[1:(n-1)])/blk$BLK.Adjusted[1:(n-1)]
# converting BlackRock returns to ts
blkRetTs <- ts(blkReturns)
plot(blkRetTs,main = "Daily Returns of BlackRock",ylab ="Returns",col = "red")
```

```{r}
# we notice that mean is around 0 and we'll check for stationarity
acf(blkRetTs,main = "ACF of daily Returns")

```

```{r}
# we notice that there are no significant spikes in ACF plot => weak dependence
pacf(blkRetTs,main = "PACF of daily Returns")
```

```{r}
# We notice that there are no significant spikes 
# Agumented Dickey Fuller Test to confrim stationrity
library(tseries)
adf.test(blkRetTs)
```
```{r}
#as p-value <0.05 we reject Ho i.e., Time series are stationary and no differencing required
size <- length(blkRetTs)
trainSize <- floor(0.8 * size)

train <- blkRetTs[1:trainSize]
test  <- blkRetTs[(trainSize + 1):size]

#fitting the modelson training set
ar1<- arima(train, order = c(1, 0, 0))
ma1 <- arima(train, order = c(0, 0, 1))
arma11 <- arima(train, order = c(1, 0, 1))
cat("AR1 fit Results:\n")
print(ar1)
cat("MA1 fit Results:\n")
print(ma1)
cat("ARMA(1,1) fit Results:\n")
print(arma11)
```
## checking residual ACF
```{r}
acf(residuals(ar1),main = "ACF of Residuals: AR(1)")
acf(residuals(ma1),main = "ACF of Residuals: MA(1)")
acf(residuals(arma11),main = "ACF of Residuals: ARMA(1,1)")
```
## We notice that all 3 are similar so we compare AIC's
```{r}
aicValues <- AIC(ar1, ma1, arma11)
cat("Lowest AIC value is:",min(aicValues$AIC),"for model:",rownames(aicValues)[order(aicValues$AIC)[1]])
```
## We noitice that phi value is near to 0 so we can forecast
```{r}
prediction <- predict(arma11, n.ahead = length(test))
fc <- prediction$pred
# 95% confidence intervals for forecasts
upper <- prediction$pred + 1.96 * prediction$se
lower <- prediction$pred - 1.96 * prediction$se

```

## Comparing forecast results with actual returns
```{r}
plot(blkRetTs,type = "l",col = "black",main = "Forecast vs Actual Returns",ylab = "Returns",xlab = "Time")

# Forecasted returns
lines((trainSize + 1):(trainSize + length(test)),fc,col = "red",lty = 2)

# Confidence intervals
lines((trainSize + 1):(trainSize + length(test)),upper,col = "blue",
      lty = 3)

lines((trainSize + 1):(trainSize + length(test)),
      lower,
      col = "blue",
      lty = 3)

# Train/test split
abline(v = trainSize, col = "darkgreen", lty = 4)

legend("topright",legend = c("Actual Returns","Forecast","95% Confidence Interval","Train/Test Split"),
       col = c("black", "red", "blue", "darkgreen"),
       lty = c(1, 2, 3, 4))


```
# Time Series Analysis of Unemployement in California State (Seasonal)
```{r}
# Import california unemployment data
caur <- read.csv("C://Users//hrush/OneDrive//Desktop//Time Series Analysis Project//CAUR.csv")

# Convert to monthly time series
cts <- ts(caur$CAUR,start = c(1976, 1),frequency = 12)

# Plot the series
plot(cts,main = "Monthly Unemployment Rate (1976–2025)",ylab = "Unemployment Rate",xlab = "Time")
```

## Checking for correlations 
```{r}
acf(cts,main = "ACF of Monthly Unemployment Rate")
```

## We notice ACF decays slowly => series are not stationary 
```{r}
cdiff <- diff(cts,differences = 1)
plot(cdiff,main ="First Differenced Unemployment rate",ylab = "Differenced Value",xlab="Time")
```
### Plot seems to be stationary after first differencing
```{r}
acf(cdiff,main = "ACF plot of differenced series ")
```

### From acf we can still see structure at lag 12(i.e., 1) so we mignt need seasonal differencing 
```{r}
cDiff <- diff(cdiff,lag = 12)
plot(cDiff,main = "Differenced Unemployment Rate (d = 1, D = 1)",ylab = "Differenced Value",xlab = "Time")
acf(cDiff,main = "ACF after Regular + Seasonal Differencing")
pacf(cDiff,main = "PACF after Regular + Seasonal Differencing")

```

```{r}
# ARIMA(1,1,0)(1,1,1)[12]
sarimaAr <- arima(cts,
                   order = c(1, 1, 0),
                   seasonal = list(order = c(1, 1, 1), period = 12))

# ARIMA(0,1,1)(1,1,1)[12]
sarimaMa <- arima(cts,
                   order = c(0, 1, 1),
                   seasonal = list(order = c(1, 1, 1), period = 12))

# ARIMA(1,1,1)(1,1,1)[12]
sarimaArma <- arima(cts,
                     order = c(1, 1, 1),
                     seasonal = list(order = c(1, 1, 1), period = 12))
cat("SARIMA(1,1,0)(1,1,1)[12] Results:\n")
print(sarimaAr)

cat("\nSARIMA(0,1,1)(1,1,1)[12] Results:\n")
print(sarimaMa)

cat("\nSARIMA(1,1,1)(1,1,1)[12] Results:\n")
print(sarimaArma)

```
## Check AIC values to choose the model
```{r}
aicValuesS <- AIC(sarimaAr,sarimaMa,sarimaArma)
cat("Lowest AIC value is:",min(aicValuesS$AIC),"for model:",rownames(aicValuesS)[order(aicValuesS$AIC)[1]])

```
## Residual Diagnostics
```{r}
acf(residuals(sarimaArma),
    main = "Residual ACF: SARIMA(1,1,1)(1,1,1)[12]")

pacf(residuals(sarimaArma),
     main = "Residual PACF: SARIMA(1,1,1)(1,1,1)[12]")

```
```{r}
Box.test(residuals(sarimaArma),lag = 24,type = "Ljung-Box")
```

```{r}
sPrediction <- predict(sarimaArma, n.ahead = 12)

sFc <- sPrediction$pred

sUpper <- sPrediction$pred + 1.96 * sPrediction$se
sLower <- sPrediction$pred - 1.96 * sPrediction$se

plot(cts,type = "l",col = "black",main = "Forecast vs Actual Unemployment Rate",ylab = "Unemployment Rate",xlab = "Time")
# Forecasted values
lines(sFc,col = "red",lty = 2)
# Confidence intervals
lines(sUpper,col = "blue",lty = 3)
lines(sLower,col = "blue",lty = 3)

legend("topright",
       legend = c("Observed","Forecast","95% Confidence Interval"),
       col = c("black", "red", "blue"),
       lty = c(1, 2, 3))

```